---
title: "Figure drafts for Jan 2017 submission"
author: "Eric Kernfeld"
date: "January 18, 2017"
output: html_document
---

## Reconsider classification of TECs

Goals for this script:

- Thoroughly describe the heterogeneity present within the TECs, accounting for uncertainty.
- Generate output describing:
    - Cluster labels to be used going forwards
    - Exploration of clustering possibilities (e.g. different numbers of clusters; dendrogram)
    - A large heatmap
    - Pie charts of cluster sizes over time
    - Marker genes and featureplots of marker genes

#### Retrieve TECs; remove doublets

```{r}
results_path = Sys.getenv( "FREEZR_DESTINATION" )

dge_atlas_TECs = inventory_get( tag = "TECS_12_p0_explore" ) %>% readRDS

save_feature_plots(dge_atlas_TECs, results_path, gene_list = c("Ptprc", "ident"), gene_list_name = "before_dub_removal")
dge_atlas_TECs %<>% SubsetDataFlex( "ident", "ident != 6" )
save_feature_plots(dge_atlas_TECs, results_path, gene_list = c("Ptprc", "ident"), gene_list_name = "after_dub_removal")

```

#### Redo exploration without doublets

```{r}
all_params = expand.grid( cc_method = c( "average" ),
                          excess_var_cutoff = c( 0.5 ),
                          log_expr_cutoff = c( 0.1 ), 
                          prop_genes_to_select = NA, 
                          num_genes_to_select = NA, 
                          var_gene_method = "seurat",
                          num_pc =  12, 
                          clust_method = "SNN",
                          clust_granularities_as_string = "0.5",
                          plot_all_var_genes = F,
                          stringsAsFactors = F)
dge_atlas_TECs %<>% explore_embeddings( results_path = results_path, all_params)
```

#### Fit cluster model with bagging

```{r}

dge_atlas_TECs %<>% add_maehrlab_metadata("eday")

NUM_CLUSTERS = 7
dge_atlas_TECs %<>% ClusterBag( k=NUM_CLUSTERS, num_pc = 25 )

# Rename clusters to be consecutive on dendrogram
{
  dge_atlas_TECs %<>% SetIdent( ident.use = dge_atlas_TECs@data.info$cl_bag_assignment )
  dge_atlas_TECs %<>% BuildClusterTree
  auto_names = extract_tip_order(dge_atlas_TECs@cluster.tree[[1]])
  phylo_names = paste0("TEC", 1:NUM_CLUSTERS)
  custom_names = c(paste0("mTEC", 1:3), 
                   paste0("cTEC", 4:1))
  old_converter = setNames(phylo_names, auto_names)
  new_converter = setNames(custom_names, auto_names)
  dge_atlas_TECs@data.info$old_cell_type  = old_converter[dge_atlas_TECs@data.info$cl_bag_assignment]
  dge_atlas_TECs@data.info$cell_type      = new_converter[dge_atlas_TECs@data.info$cl_bag_assignment]
  dge_atlas_TECs %<>% SetIdent( ident.use = new_converter[dge_atlas_TECs@data.info$cl_bag_assignment] )
}

dge_atlas_TECs %<>% BuildClusterTree
{
  pdf(file.path(results_path, "cluster_tree.pdf"), width = 6, height = 4)
  PlotClusterTree(dge_atlas_TECs)
  dev.off()
}
```

#### Summary data and cluster modeling output

```{r}

TEC_markers = c( "Pax1", "Pax9",
                 "Ccl25", "Ackr4", "Prss16", "Psmb11",  
                 "Ly75", "Enpep","Krt8", "Krt17", "Krt14",
                 "Cd74", "H2-Aa", "Aire", "Fezf2", "Tnfrsf11a", "Krt5",
                 "Cldn3", "Cldn4", "Six1", "Six4", "Eya1")

save_feature_plots(dge_atlas_TECs, results_path = results_path, types = c("PDF", "pdf_no_leg", "PDF_PNG_split"),
                   gene_list = TEC_markers, gene_list_name = "genes")
save_feature_plots(dge_atlas_TECs, results_path = results_path, types = c("PDF", "pdf_no_leg", "PDF_PNG_split"),
                   gene_list = "eday", cols.use = extended_eday_colors,                   
                   gene_list_name = "metadata") 
save_feature_plots(dge_atlas_TECs, results_path = results_path, types = c("PDF", "pdf_no_leg", "PDF_PNG_split"),
                   gene_list = "cell_type", cols.use = TEC_cols,                   
                   gene_list_name = "metadata") 
p = annot_ident_plot(dge_atlas_TECs, results_path, figname = "ident_labeled_badcol" ) + 
          scale_color_manual(values = TEC_cols)
ggsave( file.path( results_path, "ident_labeled.pdf" ), p )

save_feature_plots(dge_atlas_TECs, results_path = results_path, types = c("PDF", "pdf_no_leg", "PDF_PNG_split"),
                   gene_list =  paste0( "cl_bag_", c( "assignment", "confidence", 1:NUM_CLUSTERS ) ), 
                   gene_list_name = "metadata") 
# Check for batch/ribo effects
dge_atlas_TECs %<>% add_rp_percentage
save_feature_plots(dge_atlas_TECs, results_path = results_path, types = c("PDF", "pdf_no_leg"),
                   gene_list = c( "ribo_nUMI_pct", "orig.ident" ), 
                   gene_list_name = "metadata") 
```

#### Save labels for downstream use

```{r}
saveto = file.path( results_path, "TECS_12_P0_bag_labeled.dge" )
saveRDS(dge_atlas_TECs, saveto)
inventory_add( tag = "TECS_12_P0_bag_labeled", filename = saveto, force = T ) 
# dge_atlas_TECs = inventory_get( tag = "TECS_12_P0_bag_labeled") %>% readRDS
```

#### Appendix 1: assessing confidence for number of clusters

```{r}
conf_curve = ClusterBagStability( dge_atlas_TECs, num_pc = 25 )
conf_curve_bak = conf_curve
{
  pdf( file.path( results_path, "clustering_confidence_curve.pdf" ), width = 6, height = 3 )
  plot( main = "Bootstrap variability of clusters", 
        x = conf_curve$k, y = 1 - conf_curve$stability, 
        ylab = "Instability", xlab = "Num clusters", type = "l", ylim = c(0, 0.4) )
  dev.off()
}
```

#### Featureplots for supplement

```{r}
 
save_plot_grid( dge = dge_atlas_TECs, results_path, 
                gene_list = c( "Aire", "Krt5", "Cldn4", "H2.Aa",
                               "Ly75", "Psmb11", "Prss16" ),
                gene_list_name = "tec_supp",
                ncol = 4, 
                size = 0.5,
                leg_pos = c( 0.7, 0.95 ),
                cols.use = c("khaki1", "red"),
                width = 15, height = 7.5 )

```

