---
title: "Figure drafts for Jan 2017 submission"
author: "Eric Kernfeld"
date: "January 18, 2017"
output: html_document
---


Right now, most of the heatmapping occurs in the marker-finding script that precedes this. 
This script is for the GO term heatmaps.

#### Load data

```{r}
results_path = Sys.getenv( "FREEZR_DESTINATION" )

dge_atlas_TECs = inventory_get( tag = "TECS_12_P0_bag_labeled" ) %>% readRDS


#### Choose some genes to heatmap
de_gene_tables =
  inventory_get("TEC_subcluster_pairwise_comparisons") %>% 
  list.files(full.names = T) %>%
  lapply(read.table, header = T)
names( de_gene_tables ) = 
  inventory_get("TEC_subcluster_pairwise_comparisons") %>%
  list.files(full.names = F) %>%
  sapply(strip_suffix, ".txt")

```


#### Cell cycle scores heatmap

```{r}
cc_heat = make_heatmap_for_table( dge = dge_atlas_TECs, 
                                  genes_in_order = c( "IG1.S",
                                                      "S", 
                                                      "G2.M", 
                                                      "M",
                                                      "M.G1" ),
                                  normalize = "both", 
                                  norm_fun = standardize,
                                  desired_cluster_order = paste0("TEC", 1:7) ) 
cc_heat = cc_heat + ylab("Cell cycle score")
cc_heat = cc_heat + scale_fill_gradient2(low = "white", high = scales::muted("green"))
ggsave( file.path( results_path, "cc_heatmap_both_norm.pdf"), cc_heat )

cc_heat = make_heatmap_for_table( dge = dge_atlas_TECs, 
                                  genes_in_order = c( "IG1.S",
                                                      "S", 
                                                      "G2.M", 
                                                      "M",
                                                      "M.G1" ),
                                  normalize = "row", 
                                  norm_fun = standardize,
                                  desired_cluster_order = paste0("TEC", 1:7) ) 
cc_heat = cc_heat + ylab("Cell cycle score")
cc_heat = cc_heat + scale_fill_gradient2(low = "white", high = scales::muted("green"))
ggsave( file.path( results_path, "cc_heatmap_row_norm.pdf"), cc_heat )
```


### Heatmap some TEC-relevant genes from literature

Filter genes for detection in 10% of cells in at least one cluster.

```{r}
TEC_markers = get_TEC_genes() %>% intersect(AvailableData(dge_atlas_TECs))
pass_filter = make_heatmap_for_table(dge_atlas_TECs, 
                                     TEC_markers, 
                                     aggregator = prop_nz,
                                     normalize = "none", 
                                     return_type = "table") %>% 
  apply(1, max) %>% 
  is_greater_than(0.1) 
TEC_markers = names(which(pass_filter))
```

##### Cluster-averaged heatmap 

```{r}
dco = paste0("TEC", 1:7)
TEC_markers = make_heatmap_for_table( dge = dge_atlas_TECs, 
                                         genes_in_order = TEC_markers,
                                         ident.use = "cell_type",
                                         norm_fun = standardize,
                                         desired_cluster_order = dco,
                                         return_type = "table" ) %>% OrderRowsForHeatmap 

#### Make the actual heatmap
p = make_heatmap_for_table( dge = dge_atlas_TECs, 
                            genes_in_order = TEC_markers,
                            ident.use = "cell_type",
                            norm_fun = standardize,
                            desired_cluster_order = dco, 
                            labels = "all" ) + 
  scale_fill_gradientn( colors = blue_gray_red ) 
ggsave(file.path(results_path, "TEC_cluster_heatmap_lit.pdf"), plot=p, height = 10, width = 7)


```

##### Cellwise heatmap (many cells, few genes)

```{r}

cols_yr = colorRampPalette( c( "khaki1", "orange", "red" ) )(50)

cell_order = OrderCellsWithinClusters( dge_atlas_TECs, ident.use = "cell_type" )
norm_expr = t( apply(X = dge_atlas_TECs@data[TEC_markers, cell_order], FUN = div_by_max, MARGIN = 1) )
TEC_markers_reorder = norm_expr %>% dist %>% (fastcluster::hclust) %>% extract2("order")
TEC_markers %<>% extract(TEC_markers_reorder)
norm_expr = norm_expr[TEC_markers_reorder, ]
preview_idx = sample(3341) %>% sort
{
  pdf( file.path( results_path, "cellwise_few_genes.pdf" ), width = 27, height = 9 )
  gplots::heatmap.2( norm_expr[,preview_idx], 
                     Rowv = F, 
                     Colv = F, 
                     dendrogram = "none",
                     symm = F, 
                     scale = "none", 
                     col = cols_yr,
                     trace = "none",
                     xlab = "Cells", labCol = "",
                     ylab = "Genes", labRow = TEC_markers,
                     RowSideColors = c("black", "grey"),
                     ColSideColors = TEC_cols[ as.character(FetchData(dge_atlas_TECs, "cell_type")[cell_order[preview_idx], 1]) ] )
  dev.off()
}

try({
  rep(c(0, 1), length.out = length(TEC_markers)) %>% extract(seq_along(TEC_markers)) %>% data.frame -> X
X$fill = X$.
X$y = seq_along(X$fill)
p = ggplot(X) + 
  geom_tile(aes(fill = as.factor(fill), x = 1, y = y))  +
  scale_fill_manual(values = c("black", "grey")) -> p
ggsave("~/Desktop/colorbar.pdf", p, height = 10, width = 3)
})

```


#### Heatmap data-driven markers (small)

```{r}
genes_dd_heatmap_hi = de_gene_tables %>%
  lapply( dplyr::top_n, n = 5, wt = avg_diff ) %>%
  lapply( extract2, "gene") %>% Reduce(f=union)
genes_dd_heatmap_lo = de_gene_tables %>%
  lapply( dplyr::top_n, n = 5, wt = -avg_diff ) %>%
  lapply( extract2, "gene") %>% Reduce(f=union)
genes_dd_heatmap = union(genes_dd_heatmap_lo, genes_dd_heatmap_hi)

dco = paste0("TEC", 1:7)
genes_dd_heatmap = make_heatmap_for_table( dge = dge_atlas_TECs, 
                                         genes_in_order = genes_dd_heatmap,
                                         ident.use = "cell_type",
                                         norm_fun = standardize,
                                         desired_cluster_order = dco, 
                                         return_type = "table" ) %>% OrderRowsForHeatmap

p = make_heatmap_for_table( dge = dge_atlas_TECs, 
                            genes_in_order = genes_dd_heatmap,
                            ident.use = "cell_type",
                            norm_fun = standardize,
                            desired_cluster_order = dco, 
                            labels = "all" ) + 
  scale_fill_gradientn( colors = blue_gray_red ) 
ggsave(file.path(results_path, "TEC_cluster_heatmap_small.pdf"), plot=p, height  = 10)
```

#### Heatmap data-driven markers (large)

```{r}
#### Choose some genes to heatmap

de_gene_tables =
  inventory_get("TEC_subcluster_pairwise_comparisons") %>% 
  list.files(full.names = T) %>%
  lapply(read.table, header = T)
names( de_gene_tables ) = 
  inventory_get("TEC_subcluster_pairwise_comparisons") %>%
  list.files(full.names = F) %>%
  sapply(strip_suffix, ".txt")

TEC_markers_data_driven = de_gene_tables %>% 
  lapply(subset, avg_diff > 0.5, select = "gene", drop = T) %>% 
  lapply(as.character) %>% Reduce(f=union)

#### Choose an order for the genes
dco = paste0("TEC", 1:7)
genes_averaged = make_heatmap_for_table( dge = dge_atlas_TECs, 
                                         genes_in_order = TEC_markers_data_driven,
                                         ident.use = "cell_type",
                                         norm_fun = standardize,
                                         desired_cluster_order = dco, 
                                         return_type = "table" )

dgo = OrderRowsForHeatmap( genes_averaged )

p = make_heatmap_for_table( dge = dge_atlas_TECs, 
                            genes_in_order = dgo,
                            ident.use = "cell_type",
                            norm_fun = standardize,
                            desired_cluster_order = dco, 
                            genes_to_label = get_TEC_genes() ) + 
  scale_fill_gradientn( colors = blue_gray_red ) 
ggsave(file.path(results_path, "TEC_cluster_heatmap_big.pdf"), plot=p)

```


#### GO heatmap

```{r}

positive_markers = lapply(de_gene_tables, subset, avg_diff > 0, select = "gene", drop = T, USE.NAMES = T)
positive_markers %<>% lapply(as.character)
p = make_go_heatmap( positive_markers, effect_measure = "overlap", num_genes_tested = nrow(dge_atlas_TECs@data) ) 
p = p + theme(axis.text.x = element_text(angle = 45))
p = p + scale_fill_gradient( low = "white", high = "blue")
ggsave(file.path(results_path, "GO_heatmap_filter_lax.pdf"), p, width = 12, height = 11)


positive_markers = lapply(de_gene_tables, dplyr::top_n, n = 50, wt = avg_diff )
positive_markers %<>% lapply(extract2, "gene")
positive_markers %<>% lapply(as.character)
p = make_go_heatmap( cluster_markers = positive_markers,
                     effect_measure = "overlap", 
                     num_genes_tested = nrow(dge_atlas_TECs@data) ) 
p = p + theme( axis.text.x = element_text( angle = 45 ) )
p = p + scale_fill_gradient( low = "white", high = "blue")
ggsave(file.path(results_path, "GO_heatmap_top_50.pdf"), p, width = 12, height = 11)
```

