---
title: "Figure drafts for Jan 2017 submission"
author: "Eric Kernfeld"
date: "January 18, 2017"
output: html_document
---

## FTOC versus *in vivo* data



Assemble a big dge with ctrl and atlas.
The postconditions of this chunk are that cell_type should store the ident for the atlas 
and the classifier ident for the FTOC.
Also, edayXsource_tissue stores an interaction between eday and FTOC/in vivo.

```{r}
results_path = Sys.getenv( "FREEZR_DESTINATION" )
dge_ftoc_ctrl = inventory_get("FTOC_RA_CTRL_NO_BMS_dge") %>% readRDS
dge_ftoc_ctrl %<>% SubsetDataFlex("perturbation", "perturbation=='CTRL'"); gc()
dge_atlas = inventory_get("overview_clean_labeled") %>% readRDS
rename = plyr::rename
ct_atlas = FetchData( dge_atlas,     "ident" )            %>% rename(c("ident"="cell_type"))
ct_ctrl  = FetchData( dge_ftoc_ctrl, "classifier_ident" ) %>% rename(c("classifier_ident"="cell_type"))
ct_all = rbind(ct_atlas, ct_ctrl)
my_dge_list = list( dge_ftoc_ctrl, dge_atlas ) %>% lapply(deseuratify_raw_data)
rm(dge_atlas, dge_ftoc_ctrl); gc()
dge_atlas_plus_ctrl_ftoc = dge_merge_list(my_dge_list) 
rm(my_dge_list); gc()
dge_atlas_plus_ctrl_ftoc %<>% seuratify_thy_data(min.genes = 0, min.cells = 0)
dge_atlas_plus_ctrl_ftoc %<>% AddMetaData(ct_all, col.name = "cell_type")
dge_atlas_plus_ctrl_ftoc %<>% add_maehrlab_metadata("eday")
dge_atlas_plus_ctrl_ftoc %<>% add_maehrlab_metadata("source_tissue")
dge_atlas_plus_ctrl_ftoc %<>% add_edayXother( other = "source_tissue", other_levels = c("embryo", "FTOC"))



all_params = expand.grid( cc_method = c( "average" ),
                          excess_var_cutoff = c( 0.5 ),
                          log_expr_cutoff = c( 0.1 ), 
                          prop_genes_to_select = NA, 
                          num_genes_to_select = NA, 
                          var_gene_method = "seurat",
                          num_pc =  25, 
                          clust_method = "SNN",
                          clust_granularities_as_string = "0.5",
                          plot_all_var_genes = F,
                          stringsAsFactors = F)
dge_atlas_plus_ctrl_ftoc %<>% explore_embeddings( results_path, all_params )
dge_path = file.path( results_path, collapse_by_name(all_params[1, ]), "dge.data" )
inventory_add(tag = "dge_atlas_plus_ctrl_ftoc", 
              filename = dge_path, 
              force = T)
# dge_atlas_plus_ctrl_ftoc = inventory_get("dge_atlas_plus_ctrl_ftoc") %>% readRDS
```

##### Remove some days

We will compare the FTOC with the parallel *in vivo* development from 13.5 (FTOC extraction) to 16.5 (FTOC sequencing).

```{r}
levels_FTOC_correctly_positioned = dge_atlas_plus_ctrl_ftoc@data.info$edayXsource_tissue %>% levels %>% sort
dge_atlas_plus_ctrl_ftoc@data.info$edayXsource_tissue %<>% factor(levels = levels_FTOC_correctly_positioned )
dge_atlas_plus_ctrl_ftoc %<>% SubsetDataFlex("eday", "eday %in% (13:16 + 0.5)")
gc()
```

#### Cell type proportions 

```{r}
p = SeuratPie( dge = dge_atlas_plus_ctrl_ftoc, 
               ident.use = "cell_type", 
               facet_by = "edayXsource_tissue", 
               main = "FTOC", col = fig1_cols, drop_levels = T ) + facet_wrap("edayXsource_tissue", nrow = 2)
ggsave(file.path(results_path, "ident_by_day.pdf"), p, width = 9, height = 10)
```

#### Heatmap of selected markers

```{r}
requested_genes = c("H2-Aa", "Cd74",
                    "Ubd", "Aire", "Tbata", "Ctsl", "Prss16", "Psmb11",
                    "Foxn1", "Nfkbia", "Krt17", "Krt8","Epcam", "Il7", "Krt5",
                    "Pdgfra", "Esam",
                    "Ptprc", "Plac8",
                    "Lyz2", "Itgax", "Klrd1", "Klrb1c", 
                    "Cd8a", "Cd4", "Rag1", "Rag2",  "Ptcra","Cd3d", "Cd3e", "Cd3g", "Kit")

x = dge_atlas_plus_ctrl_ftoc %>% FetchData(c("eday", "source_tissue", "cell_type")) %>% interaction( drop = T) %>% as.character
y = dge_atlas_plus_ctrl_ftoc %>% FetchData(c("eday", "source_tissue")) %>% interaction( drop = T) %>% as.character
dge_atlas_plus_ctrl_ftoc@data.info$cell_typeXsetting = x
dge_atlas_plus_ctrl_ftoc@data.info$setting = y
{
  dco_types = paste0("C", 1:10)
  dco_source = c("embryo", "FTOC")
  dco_types  = factor(dco_types,  levels = dco_types,  ordered = T)
  dco_source = factor(dco_source, levels = dco_source, ordered = T)
  dco = levels(suppressWarnings(interaction(12:19 + 0.5, dco_source, dco_types)))
}
p = make_heatmap_for_table( dge = dge_atlas_plus_ctrl_ftoc, 
                            genes_in_order = requested_genes,
                            desired_cluster_order = dco, 
                            ident.use = "cell_typeXsetting", 
                            labels = "all")
ggsave(file.path(results_path, "main_markers_heatmap.pdf"), p, width = 24, height = 12)
```

#### Thymocyte maturation 

Subset thymocytes

```{r}
dge_atlas_plus_ctrl_ftoc %<>% SubsetDataFlex("cell_type", "cell_type %in% paste0('BLD', 1:3)")
gc()
```

via TACS

```{r}
cutoff_CD25 = 1
cutoff_CD44 = 0.7
dge_atlas_plus_ctrl_ftoc@data.info$edayXsource_tissue %<>% 
  factor(levels = sort(levels(dge_atlas_plus_ctrl_ftoc@data.info$edayXsource_tissue)))
dge_atlas_plus_ctrl_ftoc %<>%
  TACS(gene1 = "Il2ra", gene2 = "Cd44", cutoffs = c( cutoff_CD25, cutoff_CD44 ),
       facet_by = "edayXsource_tissue", return_val = "seurat") 
p_TACS_bld = dge_atlas_plus_ctrl_ftoc %>%
  TACS(gene1 = "Il2ra", gene2 = "Cd44", cutoffs = c( cutoff_CD25, cutoff_CD44 ),
       facet_by = "edayXsource_tissue", density = T) 
p_TACS_bld = p_TACS_bld + facet_wrap("edayXsource_tissue", nrow = 1)
p_TACS_bld = p_TACS_bld + scale_color_manual(values = extended_eday_ftoc_colors )
ggsave(file.path(results_path, "p_TACS_bld.pdf"), p_TACS_bld, height = 4, width = 15)
X = dge_atlas_plus_ctrl_ftoc %>% FetchData(c("edayXsource_tissue", "Il2ra_score", "Cd44_score"))

X$Il2ra_score %<>% is_greater_than(cutoff_CD25)
X$Cd44_score %<>% is_greater_than(cutoff_CD44)
colnames(X)[2:3] = c("Il2ra_hi", "Cd44_hi")
X_counts = table(X) %>% apply(1, c) %>% extract(, c("13.5.embryo", "16.5.FTOC"))
{
  sink(file.path(results_path, "fisher_test_tcell.txt"))
  fisher.test( X_counts, hybrid = T ) %>% print
  print("Cell count:")
  X_counts %>% sum
  sink()
}
write.table(table(X) %>% as.data.frame(), file.path(results_path, "Tcell_counts.txt"), quote = F)
write.table(X_counts, file.path(results_path, "Tcell_counts_tested.txt"), quote = F)

```

#### Thymocyte maturation 

via fractional identities 

```{r}

eefc = extended_eday_ftoc_colors
names(eefc) %<>% make.names
fi_out = display_fractional_identity( dge_test  = dge_atlas_plus_ctrl_ftoc, 
                                      dge_train = dge_atlas_plus_ctrl_ftoc %>%
                                        SubsetDataFlex( "source_tissue", "source_tissue=='embryo'" ),
                                      main = "FTOC maturity (thymocytes only)",
                                      ident.use = "eday",
                                      test.ident.use = "edayXsource_tissue",
                                      test.col = eefc,
                                      genes.use = NULL )


ggsave( file.path( results_path, "fractional_identities_thymocyte_FTOC_identities_hidden.pdf" ), 
        fi_out$plot,
        width = 8, height = 2.5 )
```


