---
title: "Figure drafts for Jan 2017 submission"
author: "Eric Kernfeld"
date: "January 18, 2017"
output: html_document
---

## FTOC versus *in vivo* TECs

Assemble a big dge with ctrl and atlas.
The postconditions of this chunk are that cell_type should store the ident for the atlas 
and the classifier ident for the FTOC.
Also, edayXsource_tissue stores an interaction between eday and FTOC/in vivo.

```{r}
results_path = Sys.getenv( "FREEZR_DESTINATION" )

dge_atlas_ftoc_ctrl_TEC = inventory_get( tag = "dge_FTOC_RA_CTRL_NO_BMS_tec" ) %>% readRDS %>% 
  SubsetDataFlex( "perturbation", "perturbation=='CTRL'")
dge_atlas_ftoc_ctrl_TEC %<>% SeuratMerge( inventory_get("TECS_12_P0_bag_labeled") %>% readRDS )
dge_atlas_ftoc_ctrl_TEC %<>% add_edayXother( other = "source_tissue", other_levels = c("embryo", "FTOC"))

all_params = expand.grid( cc_method = c( "average" ),
                          excess_var_cutoff = c( 0.5 ),
                          log_expr_cutoff = c( 0.1 ), 
                          prop_genes_to_select = NA, 
                          num_genes_to_select = NA, 
                          var_gene_method = "seurat",
                          num_pc =  25, 
                          clust_method = "SNN",
                          clust_granularities_as_string = "0.5",
                          plot_all_var_genes = F,
                          stringsAsFactors = F)
dge_atlas_ftoc_ctrl_TEC %<>% explore_embeddings( results_path, all_params)
dge_path = file.path( results_path, collapse_by_name(all_params[1, ]), "dge.data" )
inventory_add(tag = "dge_ftoc_ctrl_plus_atlas_TEC", 
              filename = dge_path, 
              force = T)
# dge_atlas_ftoc_ctrl_TEC = inventory_get( "dge_ftoc_ctrl_plus_atlas_TEC" ) %>% readRDS

```

Contextual info

```{r}

dge_atlas_ftoc_ctrl_TEC %<>% add_maehrlab_metadata("source_tissue")
dge_atlas_ftoc_ctrl_TEC %>% save_feature_plots(results_path,
                                types = c("PDF", "PDF_no_leg", "PNG_PDF_split"),
                                gene_list = "source_tissue",
                                gene_list_name = "metadata")

dge_atlas_ftoc_ctrl_TEC %>% save_feature_plots(results_path,
                                cols.use = TEC_cols, 
                                types = c("PDF", "PDF_no_leg", "PNG_PDF_split"),
                                gene_list = "cell_type",
                                gene_list_name = "metadata")

dge_atlas_ftoc_ctrl_TEC %>% save_feature_plots(results_path,
                                types = c("PDF", "PDF_no_leg", "PNG_PDF_split"),
                                gene_list = "plain_grey",
                                gene_list_name = "metadata")

dge_atlas_ftoc_ctrl_TEC %>% save_feature_plots(results_path,
                                               cols.use = extended_eday_colors,
                                               types = c("PDF", "PDF_no_leg", "PNG_PDF_split"),
                                               gene_list = "eday",
                                               gene_list_name = "metadata")

dge_atlas_ftoc_ctrl_TEC %>% save_feature_plots(results_path,
                                               cols.use = extended_eday_ftoc_colors,
                                               types = c("PDF", "PDF_no_leg", "PNG_PDF_split"),
                                               gene_list = "edayXsource_tissue",
                                               gene_list_name = "metadata")
```

Cell type proportions 

```{r}
dge_atlas_ftoc_ctrl_TEC@data.info$cell_type = "initialize"
TEC_assignments_atlas = inventory_get("TECS_12_P0_bag_labeled") %>% readRDS %>% FetchData("cell_type")
TEC_assignments_ftoc = knn_classifier(dge_train = inventory_get("TECS_12_P0_bag_labeled") %>% readRDS, 
                                      dge_test  = inventory_get("dge_FTOC_RA_CTRL_NO_BMS_tec" ) %>% readRDS, 
                                      ident.use = "cell_type") 
TEC_assignments_ftoc %<>% FetchData("classifier_ident")
dge_atlas_ftoc_ctrl_TEC@data.info[rownames(TEC_assignments_atlas), "cell_type"] = TEC_assignments_atlas[, 1] %>% as.character
dge_atlas_ftoc_ctrl_TEC@data.info[rownames(TEC_assignments_ftoc),  "cell_type"] =  TEC_assignments_ftoc[, 1] %>% as.character

dge_atlas_ftoc_ctrl_TEC %<>% SubsetDataFlex("eday" , "eday %in% (13:16 + 0.5)")
dge_atlas_ftoc_ctrl_TEC %<>% SanitizeGenes
p = SeuratPie( dge = dge_atlas_ftoc_ctrl_TEC, 
               ident.use = "cell_type", 
               facet_by = "edayXsource_tissue", 
               main = "FTOC", col = TEC_cols, drop_levels = T ) +
  facet_wrap("edayXsource_tissue", nrow = 2 )
ggsave(file.path(results_path, "ident_by_day.pdf"), p, width = 15, height = 5)

{
  sink(file.path(results_path, "fisher_test_TEC_subtypes_16_5_v_FTOC.pdf"))
  dge_atlas_ftoc_ctrl_TEC %>% FetchData(c("edayXsource_tissue", "cell_type")) %>% table %>% extract(c("16.5.FTOC", "16.5.embryo"), ) %>% as.matrix %>% fisher.test(hybrid = T)
  print("Cell count:")
  dge_atlas_ftoc_ctrl_TEC %>% FetchData(c("edayXsource_tissue", "cell_type")) %>% table %>% extract(c("16.5.FTOC", "16.5.embryo"), ) %>% as.matrix %>% sum
  sink()
}

```

TACS to measure overall maturity

```{r}
cutoff_H2.Aa  = 2
cutoff_Epcam = 1.5
p_TACS_tec = dge_atlas_ftoc_ctrl_TEC %>%
  TACS(gene1 = "H2.Aa", gene2 = "Epcam", cutoffs = c( cutoff_H2.Aa, cutoff_Epcam ),
       facet_by = "edayXsource_tissue",  density = T) + 
  scale_color_manual(values = extended_eday_ftoc_colors) + 
  facet_wrap("edayXsource_tissue", nrow = 1)
ggsave(file.path(results_path, "p_TACS_tec_maturation.pdf"), p_TACS_tec, width = 15, height = 4)

dge_atlas_ftoc_ctrl_TEC %<>% TACS(gene1 = "H2.Aa", gene2 = "Epcam", 
                                  cutoffs = c( cutoff_H2.Aa, cutoff_Epcam ),
                                  facet_by = "edayXsource_tissue",  return_val = "seurat") 
X = dge_atlas_ftoc_ctrl_TEC %>% FetchData(c("edayXsource_tissue", "H2.Aa_score", "Epcam_score"))
X$H2.Aa_score %<>% is_greater_than(cutoff_H2.Aa)
X$Epcam_score %<>% is_greater_than(cutoff_Epcam)
colnames(X)[2:3] = c("H2.Aa_hi", "Epcam_hi")
X_counts = table(X) %>% apply(1, c) %>% extract(, c("13.5.embryo", "16.5.FTOC"))
{
  sink(file.path(results_path, "fisher_test_TEC_ftoc.txt"))
  fisher.test( X_counts, hybrid = T ) %>% print
  print( "Cell counts" )
  sum(X_counts)
  sink()
}
write.table(table(X) %>% as.data.frame(), file.path(results_path, "TEC_counts.txt"), quote = F)

```

Fractional identity to measure maturity

```{r}
eefc = extended_eday_ftoc_colors
names(eefc) %<>% make.names
fi_out = display_fractional_identity( dge_test  = dge_atlas_ftoc_ctrl_TEC, 
                                      dge_train = dge_atlas_ftoc_ctrl_TEC %>%
                                        SubsetDataFlex( "source_tissue", "source_tissue=='embryo'" ),
                                      main = "FTOC maturity (TECs only)",
                                      ident.use = "eday",
                                      test.ident.use = "edayXsource_tissue",
                                      test.col = eefc,
                                      genes.use = NULL )


ggsave( file.path( results_path, "fractional_identities_TEC_FTOC_identities_hidden.pdf" ), 
        fi_out$plot,
        width = 8, height = 2.5 )

```







