---
title: "Figures for Thymic Development Atlas"
author: "Eric Kernfeld"
date: "February 17, 2017"
output: html_document
---

### Setup

This tells the code where to find itself. 

```{r setup, include=FALSE}

# You may need to run setwd(proj_dir) in the console due to knitr's well-documented working directory insanity.
proj_dir = "~/Desktop/thymusatlasanalysis"
setwd(file.path(proj_dir, "scripts"))
```

You will need to install these packages.

```{r}
PATH_TO_PROJECT_PACKAGES = file.path(proj_dir, "R_packages")
dir.create(PATH_TO_PROJECT_PACKAGES)
.libPaths(PATH_TO_PROJECT_PACKAGES)
withr::with_libpaths( PATH_TO_PROJECT_PACKAGES, devtools::install_github( "satijalab/seurat", ref = "d9d1a8b" ) )
withr::with_libpaths( PATH_TO_PROJECT_PACKAGES, devtools::install_github( "maehrlab/thymusatlastools" ) )
withr::with_libpaths( PATH_TO_PROJECT_PACKAGES, devtools::install_github( "ekernf01/freezr" ) )
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
BiocManager::install("destiny", version = "3.8" )
library(Seurat,           lib.loc = PATH_TO_PROJECT_PACKAGES)
library(thymusatlastools, lib.loc = PATH_TO_PROJECT_PACKAGES)
library(freezr,           lib.loc = PATH_TO_PROJECT_PACKAGES)
```

This sets up the workspace.

```{r}
get_metadata_path = function() file.path( proj_dir, "data", "metadata.csv")
get_metadata = function() read.csv(file.path(get_metadata_path()), stringsAsFactors = F)

#' This is specific to my folder layout. 
#' When run without args, this function runs only thymus_functions.Rmd and saves to the "interactive" folder.
#' Otherwise, it runs both thymus_functions.Rmd and an analysis script, and it 
#' names the results subfolder after the source code subfolder. This way, the 
#' organizational structure is mirrored between the analysis scripts and the results folders.
#' 
today_dest = file.path( proj_dir, "results" )
flash_freeze = function( analysis_to_run = NULL, results_subdir = NULL ){
  if(is.null(results_subdir)){
     if(is.null(analysis_to_run)){
      results_subdir = "interactive"
    } else {
      results_subdir = dirname(analysis_to_run)
    }
  }
  freezr::freeze( analyses_to_run = c( "reusable/thymus_functions.Rmd", analysis_to_run ),
                  destination = file.path(today_dest, results_subdir),
                  chastise = F, seed_to_set = 20170322, 
                  repos_to_track = NULL )
}
```

#### Inventory setup

This bit of code creates an inventory, which is `freezr`'s way of tracking important data files. The first file tracked is a table to help convert between two sets of cluster names that were both considered during the project. 

```{r}
atlas_inventory_location = file.path(proj_dir, "results")
inventory_exists( inv_location = atlas_inventory_location )
try( inventory_make(   inv_location = atlas_inventory_location ) )
inventory_add( tag = "overview_old_labels",
               filename = file.path(proj_dir, "tables", "fig1_label_conversions.txt" ), 
               inv_location = atlas_inventory_location )
```

#### Namespace setup

This interactive run loads in `thymus_functions.Rmd` so you can play around using our functions and color scales.

```{r}
flash_freeze()
```

### Data assembly and initial exploration

```{r}
last_dest = flash_freeze( "overview/draft_data_assemble.Rmd" )
last_dest = flash_freeze( "overview/draft_doublet_fig.Rmd" )
```

### Exploration of clean data

```{r}
last_dest = flash_freeze( "overview/draft_do_clean_tsne.Rmd" )
last_dest = flash_freeze( "overview/draft_relabel_clean_tsne.Rmd" )
last_dest = flash_freeze( "overview/draft_get_markers.Rmd" )
last_dest = flash_freeze( "overview/draft_overview_heatmaps.Rmd" )
last_dest = flash_freeze( "overview/draft_overview_make_depict_input.Rmd" )
```

### Describe the ConvT and NCL compartments

```{r}
last_dest = flash_freeze( "overview/draft_overview_Tcell.Rmd" )
last_dest = flash_freeze( "NCL/draft_NCL_explore.Rmd" ) 
last_dest = flash_freeze( "NCL/draft_NCL_markers.Rmd" ) 
last_dest = flash_freeze( "NCL/draft_NCL_lit_tables.Rmd" ) 
last_dest = flash_freeze( "NCL/draft_NCL_make_depict_input.Rmd" ) 
```

### Describing the atlas TEC heterogeneity

Set up mTEC/cTEC labels. Add 16.5 sorted data and explore/relabel the result.

```{r}
last_dest = flash_freeze( "TEC_subsets/draft_TEC_subset_12_P0.Rmd" )
last_dest = flash_freeze( "TEC_subsets/draft_TEC_het_concise_12_P0.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_TEC_het_markers.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_TEC_het_heatmaps.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_TEC_het_atlas_plus_sorted.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_TEC_het_atlas_plus_adult.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_cTEC_pt.Rmd" ) 
last_dest = flash_freeze( "TEC_subsets/draft_mTEC_pt.Rmd" ) 
```

## Exploring perturbations and Rag1 knockouts

#### Rag1 knockout

These scripts:

- perform overall exploration of the Rag1 KO data (`_explore`), 
- establish that the T-cells are strongly affected (`_Tcell`),
- explore possible effects on the TECs (`_TECs`),
- explore possible effects on the non-conventional T cells (`_NKT`).

The initial exploratory script performs the T-cell isolation for the Theis lab.

```{r}
last_dest = flash_freeze( "KO/draft_Rag1_knockout_explore.Rmd" ) 
last_dest = flash_freeze( "KO/draft_Rag1_knockout_Tcell.Rmd" ) 
last_dest = flash_freeze( "KO/draft_Rag1_knockout_NCL.Rmd" ) 
last_dest = flash_freeze( "KO/draft_Rag1_knockout_TEC_explore.Rmd" ) 
last_dest = flash_freeze( "KO/draft_Rag1_knockout_TECs.Rmd" ) 
```

#### FTOC 

```{r}
last_dest = flash_freeze( "FTOC_RA/FTOC_explore.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_explore_control.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_assess_control.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_assess_control_TEC.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_TEC.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_Tcell.Rmd" ) 
last_dest = flash_freeze( "FTOC_RA/FTOC_NCL.Rmd" ) 
```

###Odds and ends

```{r}
# # Do some extra QC, checking replicates via t-SNE after removal of doublets
last_dest = flash_freeze( "qc/draft_extra_qc.Rmd" )
# # QC on peripheral data 
last_dest = flash_freeze( "draft_peripheral_qc.Rmd" )
```

